---
title: "The Central Limit Theorem and the Exponential Distribution "
author: "Garoe Gonzalez"
date: "24 de octubre de 2015"
output: pdf_document
---
## Introduction

In this project we use R to investigate the exponential distribution and compare it with the Central Limit Theorem (CLT).

In probability theory and statistics, the exponential distribution is the probability distribution that describes the time between events from a process in which events occur continuously and independently at a constant average rate. The mean of the exponential distribution is $\mu=1/\lambda$ and the standard deviation is also $\sigma=1/\lambda$.

In this project we probe the the central limit theorem using an exponential distribution. This theorem states that, given certain conditions, the arithmetic mean of a sufficiently large number of iterates of independent random variables, each with a well-defined expected value and well-defined variance, will be approximately normally distributed, regardless of the underlying distribution.

## Expontential distributions simulation 

We want to investigate how the arithmetic mean from exponentially distributed samples converges to a normal distribution (proving the CLT). This distribution can be simulated in R with $rexp(n, 1/\lambda)$. For all of the simulations, we set $\lambda=0.2$, and we investigate the distribution of averages of 40 exponentials. We will consider 1000 simulations.

The following code simulates 1000 exponentially distributed samples with 40 values each:

```{r}
set.seed(1)     # for reproducibility
n_sim  <- 1000  # number of simulated averages
lambda <- 0.2   # rate parameter
#simulation
mns = NULL
for (i in 1 : n_sim) mns = c(mns, mean(rexp(40,rate = 1/lambda)))
```

## Results
### Simulated vs theoretical mean
From the 1000 simulations we compare their expected mean with the theoretical one which by construction is $\mu = \lambda =0.2$:
```{r}
mean(mns)
```
We observed how the results is indeed close the is theoretical value.
In order to illustrate this, the distribution is plotted below as a density plot, where the position of the sample mean is indicated by a vertical blue line.

```{r}
library(ggplot2)
ggplot(data.frame(mns), aes(x = mns)) + 
     geom_density() +
     geom_vline(xintercept=mean(mns), size = 1, color = 'blue') 
````

### Simulated vs theoretical variance

The theoretical value for the variance of the distribution of averages is given by the variance of the original population $\sigma^2$ divided by the number of samples $n$ used to compute the averages : $var(sample)=\frac{\sigma^2}{n}=\frac{\lambda^2}{40}=0.001$. If we compute the variance of the samples, we conclude it is very close to the expected one:

```{r}
var(mns)
```

### Probing the Central Limit Theorem

In order to compare the distribution of averages to the Central Limit Theorem (CLT), we plot below the distribution as an histogram along with a normal distribution N($\mu$,$\sigma/\sqrt{n}$).

```{r}
mu    <- lambda
sigma <- lambda
ggplot(data.frame(mns), aes(x = mns)) + 
geom_histogram(alpha = .10, binwidth=0.01, colour = "black", aes(y = ..density..)) +
stat_function(geom = "line", fun = dnorm, arg = list(mean = mu, sd = sigma/sqrt(40)),
              size = 2, color = "blue", fill = NA)
```
